NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-04-13 Manuscript/manuscriptPS.tex
Update manuscriptPS.tex
Edits to the abstract

\begin{abstract}
  % insert abstract here
Phosphatidylserine (PS) is a negatively
charged lipid commonly found in eukaryotic membranes, where it interacts with signaling and other proteins via
electrostatic interactions and direct binding, and can induce
membrane fusion and phase separation in the presence of calcium ions.
Molecular details of these phenomena are not well understood,
because accurate models to interpret the experimental data have not
been available. Here, we use solid-state NMR S-DROSS experiments to measure the signs of the previous set of experimental NMR
order parameter magnitudes available for POPS to assess the phosphoserine structure in the different atomistic force-fields presently used.    
We also gather a set of experimental NMR data that can be used to assess the ion binding to pure PS and mixed PC:PS lipid bilayers
in molecular dynamics (MD) simulations.
Using the open collaboration
method, we extract data from wide range of available PS MD models (force fields) and perform a comparison to NMR results.
%or: Using the open collaboration method, we collect a database of MD simulations using wide range of available PS models (force fields) and perform a comparison to NMR results
We find that none of the models reproduce the NMR data within experimental accuracy,
but the best ones suggest that the carboxyl group in the serine headgroup
does not rotate freely. In line with the previous results for PC lipids,
none of the PS force fields correctly captures the cation binding affinity.
Furthermore, the response of PS headgroups to bound ions
\emph{qualitatively} differs from experiments. The collected experimental dataset and simulation
results pave the way for improvement of lipid force fields to correctly
describe the biologically relevant negatively charged membranes and their interactions with ions.
This work is part of the NMRlipids open collaboration project (\url{nmrlipids.blogspot.fi}).
\end{abstract}

>>>>

\begin{abstract}
  % insert abstract here
Phosphatidylserine (PS) is a negatively
charged lipid commonly found in eukaryotic membranes, where it interacts with proteins via
electrostatic interactions and direct binding, and can induce
membrane fusion and phase separation in the presence of calcium ions.
Molecular details of these phenomena are not well understood,
because accurate models to interpret the experimental data have not
been available. Here, we use solid-state NMR S-DROSS to measure the signs of the experimental POPS headgroup
order parameters, and gather a set of experimental NMR results from pure PS and mixed PC:PS lipid bilayers. Using the open collaboration
method, we then extract data from a wide range of available PS MD models (force fields) and asses the PS headgroup structures and the ion binding produced by these force fields by comparing to the NMR data. We find that none of the models reproduce the NMR data within experimental accuracy,
but the best ones suggest that the carboxyl group in the serine headgroup
does not rotate freely. In line with the previous results for PC lipids,
none of the PS force fields correctly captures the cation binding affinity---the response of PS headgroups to bound ions can even
\emph{qualitatively} differ from experiments. The collected experimental dataset and simulation
results pave the way for improvement of lipid force fields to correctly
describe the biologically relevant negatively charged membranes and their interactions with ions.
This work is part of the NMRlipids open collaboration project (\url{nmrlipids.blogspot.fi}).
\end{abstract}
